---
title: "FinalProject"
author: "Justin Mao"
date: "4/13/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Clean the data}
speed <- read.csv("Speed Dating Data.csv")

speed$condtn <- as.factor(speed$condtn)
speed$gender <- as.factor(speed$gender)
speed$match <- as.factor(speed$match)
speed$field_cd <- as.factor(speed$field_cd)
levels(speed$field_cd) <- c("Law","Math","SocScie/Psych", "MedSci", "Engineering", "English", "History", "Business", "Education", "Bio","SocialWork","Undergrad", "PoliSci", "Film","FineArts","Lang","Architecture","Other")
speed$race <- as.factor(speed$race)
speed$goal <- as.factor(speed$goal)
levels(speed$goal) <- c("FunNightOut", "MeetNewPpl", "GetADate","SRSRelationship", "ToSayIDidIt","Other")
speed$date <- as.factor(speed$date)
levels(speed$date) <- c("SVRL/Week","2/Week","1/Week","2/Month", "1/Month", "SVRL/Year", "AlmostNever")
speed$go_out <- as.factor(speed$go_out)
levels(speed$go_out) <- c("SVRL/Week","2/Week","1/Week","2/Month", "1/Month", "SVRL/Year", "AlmostNever")
speed$career_c <-as.factor(speed$career_c)
levels(speed$career_c) <- c("Lawyer","Academic/Research","Psychologist","DocMed", "Engineer", "Entertainment", "Banking/Consulting", "RealEstate","IntlAffairs","Undecided","SocialWork","SpeechPath","Politics", "ProSports", "Other", "Journalism", "Architecture")
speed$race_o <-as.factor(speed$race_o) 
speed$dec_o <- as.factor(speed$dec_o)
speed$samerace <- as.factor(speed$samerace)

sd2 <- speed
sd2 <- sd2[ , -1] #IID  
sd2 <- sd2[, -1] #ID  
sd2 <- sd2[, -2] #IDG
sd2 <- sd2[, -3] #Wave
sd2 <- sd2[, -3] #Round
sd2 <- sd2[, -3] #Position
sd2 <- sd2[, -3] #Postion1
sd2 <- sd2[, -4] #Partner 
sd2 <- sd2[, -4] #PID
sd2 <- sd2[, -26]#Field
sd2 <- sd2[, -(27:29)]#Academics
sd2 <- sd2[,-(30:32)]#Socioeconomic 
sd2 <- sd2[,-33]#Career
sd2 <- sd2[,-(59:64)]#What others look for
sd2 <- sd2[,-(70:74)]#Others perception
sd2 <- sd2[,-(81:92)]#Data gathered after intitial
sd2 <- sd2[,(1:79)]
sd2 <- sd2[,-(70:79)] #Removes Post First Date
sd2 <- sd2[,-52]#exclude expnum

sdrandom <- sd2[sample(nrow(sd2), nrow(sd2)),] #Get a random sample since the data is organized by participant
sdclean <- na.omit(sdrandom) #Remove rows with NA values to create a "clean" set




library(ggplot2)
library(class)
library(gmodels)
library(zoo)
library(lmtest)
library(car)
library(dplyr)
library(kernlab)
library(neuralnet)
library(caret)


sdclean$gender <- factor(x=sdclean$gender, levels = c(0,1), labels = c("Female","Male"))
sdclean$race <-factor(x=sdclean$race, levels = c(1,2,3,4,5,6), labels = c("Black/African American","European/Caucasian-American","Latino/Hispanic American","Asian/Pacific Islander/Asian-American","Native American","Other"))
sdclean$race_o <- factor(x=sdclean$race_o, levels =c(1,2,3,4,5,6), labels = c("Black/African American","European/Caucasian-American","Latino/Hispanic American","Asian/Pacific Islander/Asian-American","Native American","Other"))

```

```{r GGplot for race}
sdclean_new <- sdclean %>% group_by(gender, race) %>% summarise(Count = n())
View(sdclean_new)
ggplot(sdclean_new, aes(x=factor(gender), y=Count, fill=factor(race))) + 
    geom_bar(stat="identity") + theme_bw() + ggtitle("Demographic Distribution") + labs(x="Gender",y="Number")+theme(plot.title = element_text(size=20, face="bold", margin = margin(10, 0, 10, 0)))+theme(legend.title=element_blank())
```

```{r GGplot for career}
sdclean_career <- sdclean %>% group_by(gender, career_c) %>% summarise(Count = n())
View(sdclean_career)
ggplot(sdclean_career, aes(x=factor(gender), y=Count, fill=factor(career_c))) + 
    geom_bar(stat="identity") + theme_bw() + ggtitle("Occupation Distribution") +theme(plot.title = element_text(size=20, face="bold", 
    margin = margin(10, 0, 10, 0))) +labs(x="Gender", y="Number")+theme(legend.title=element_blank())
```

####Conclusion:  About 1/3 of female are from academic/research fields and alomost half of male are working on banking/consulting, followed by a big chunk of academic/research.

```{r Question 6 Analysis in terms of race}
sdclean$race_comparison <- ifelse(sdclean$race==sdclean$race_o,"1","0")
sdclean_race_comparison <- sdclean %>% group_by(match, race_comparison) %>% summarise(Count = n())
sdclean_race_comparison$match<- factor(x=sdclean_race_comparison$match, levels = c(0,1), labels = c("Not Match","Match"))
sdclean_race_comparison$race_comparison<- factor(x=sdclean_race_comparison$race_comparison, levels = c(0,1), labels = c("Different Race","Same Race"))
View(sdclean_race_comparison)
ggplot(sdclean_race_comparison, aes(x=factor(race_comparison), y=Count, fill=factor(match))) + 
    geom_bar(stat="identity") + theme_bw() + ggtitle("Match Distribution") +theme(plot.title = element_text(size=20, face="bold", 
    margin = margin(10, 0, 10, 0))) +labs(x="", y="")+theme(legend.title=element_blank())
```

####Conclusion: This chart doesn't show significant difference between match/not match percentage in both "different race" and "same race" category. Further analysis reaffirms that "match"percentage in "different race" and "same race" is 17.0% and 17.6%, respectively. Therefore, race might not be a significant variable to predict the match.

```{r Question 6 Further Analysis}
sdclean_race_imprace <- sdclean %>% group_by(match) %>% summarise(avg = mean(imprace))
sdclean_race_imprace

```

####Conclusion: Given the scale (1-10) indicating the importance of dating with someone who has the same race, the average points of this scale for people end up match verus not match is 3.48 and 3.84, respectively, which means what people state what they want before the event is  different from what actually influences their decisions. In this case, race has little influence on thier date choices. 

```{r}


```